---
title: "Initial Korean covid 19"
author: "Gracie Telles"
date: "Fall 2023"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well  

```{r}
str(kRdata)
```

# Your Question - Is dying independent of sex?
The question is, is the dying independent of sex? 

$H_0$: Null hypothesis is there is no relationship between the sex of the individual and their death.

$H_A$: Alternative hypothesis is there is a relationship between an individual's sex and their death.


## Method
The response variable is categorical, and the explanatary variable is also categorical, so we use the cat-cat helper file.


## Graphical results - use ggplot commands to do this


```{r}
dd2 <- kRdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))

basicC <- ggplot(dd2,aes(x=Sex,y=count,fill=result))
basicC + geom_bar(stat="identity",position="dodge")


#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=Sex,y=prcnt*100,fill=result)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

The bar graph shows us that the percentage of men who died from Covid 19 was overwhelmingly higher than that of women.


  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

When you compare the percentage of men versus women who died from Covid 19, men died twice as much as women. 64% of men died, and only 36% of women died from this virus.

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

In both the Fisher and Chi-Squared tests, the P value is less than 0.05. Therefore we reject the null hypothesis.

## Conclusions

The original question was if dying of Covid 19 was independent of a person's sex. Our graphical and numerical data shows that male deaths were about two times higher than women's, and because the P value of the data was less than 0.05, we reject the null hypothesis. So males are more likely than females to die due to Covid 19.